---
title: "ADA Final Project"
author: "Meiqi Liu"
date: "2022-12-01"
output: html_document
---

# Load and library packages
```{r}
pacman::p_load(haven, dplyr, odds.n.ends, blorr, lmtest, car,  broom, tidyverse, jtools, table1, DiagrammeR) 
```

# Import sav dataset
```{r}
HINTS <- read_sav("D:/wustl/ADAProject/hints5_cycle4_public.sav")
```

# Select variables
```{r}
# get head of dataset
head(HINTS)

# select variables we need
HINTS0 <- HINTS %>% 
  select(2,286,314,315,317,424,427)
```

# Rename variables
```{r}
HINTS0<-HINTS0 %>% 
  rename("ID"=1, "Sex"=4, "Employ"=5, "Educ"=6, "RaceEth"=7)
```

# Target population: women in 40-65
```{r}
# age: 40-65
HINT4065<-subset(HINTS0, HINTS0$Age>39 & HINTS0$Age<66)

# sex: female
FEMALE4065<-subset(HINT4065, HINT4065$Sex==2)
```

# Summary dataset
```{r}
summary(FEMALE4065)
```
# Recode Sex
```{r}
FEMALE4065$Sex<-factor(FEMALE4065$Sex, 2:2, c("Female"))
```

# Recode Breast cancer
```{r}
FEMALE4065<-FEMALE4065 %>% 
  mutate(BreastCan = case_when(CaBreast==1~1,
                               CaBreast==2~0),
         BreastCan = factor(BreastCan, 0:1, c("Not Diagnosed Breast Cancer", "Diagnosed Breast Cancer")))

table(FEMALE4065$CaBreast, FEMALE4065$BreastCan, useNA = "always")
```

# Recode Employment
```{r}
FEMALE4065<-FEMALE4065 %>% 
  mutate(Employment = case_when(Employ==2~0,
                                Employ==1~1),
         Employment = factor(Employment, 0:1, c("No","Yes")))

table(FEMALE4065$Employ,FEMALE4065$Employment, useNA = "always")
```

# Recode Educ
```{r}
FEMALE4065<-FEMALE4065 %>% 
  mutate(Educ = na_if(Educ,-9),
         Educ = factor(Educ, 1:4, c("Less than High school", "High school Graduate", "Some College", "College Graduate or more")))
```

# Recode Race/Ethnicity
```{r}
FEMALE4065<-FEMALE4065 %>% 
  mutate(RaceEthnicity = na_if(RaceEth, -9),
         RaceEthnicity = factor(RaceEthnicity, 1:5, c("Non-Hispanic White", "Non-Hispanic Black or African American", "Hispanic","Non-Hispanic Asian","Non-Hispanic Other")))
```

# Drop NAs
```{r}
HINTS1<-FEMALE4065 %>% 
  select(1,3,4,6,8,9,10) %>% 
  drop_na()
```

# Summary dataset
```{r}
summary(HINTS1)
```

# Figure 1: shows exclusions to arrive at the analytic dataset
```{r}
grViz("digraph flowchart {

      node [fontname = Helvetica, shape = rectangle, fontsize=10] 
      
      tab1 [label = '@@1'] # starting
      tab2 [label = '@@2'] # exclusion 1
      tab3 [label = '@@3'] # exclusion 2
      tab4 [label = '@@4'] # exclusion 3
      
      tab1 -> tab2 -> tab3 -> tab4
}
     
      [1]: 'Records received from HINTS n=3,865'
      [2]: 'Excluding 2,122 individuals not aged in 40-65 n=1,743'
      [3]: 'Excluding 721 male individuals n=1,022'
      [4]: 'Excluding 889 individuals with missing data on variables used in the analysis n=133'
      ")
```

# table 1
```{r}
label(HINTS1$Age)<-"Age (years)"
label(HINTS1$Sex)<-"Sex"
label(HINTS1$Educ)<-"Educational Level"
label(HINTS1$BreastCan)<-"Breast Cancer Diagnosis"
label(HINTS1$RaceEthnicity)<-"Race/Ethnicity"
table1(~Age+Sex+Educ++RaceEthnicity+Employment|BreastCan, overall="Total", rowlabelhead="Variable", footnote='SD = standard deviation', HINTS1)
```

# Regroup strata have <10 counts
```{r}
# education
HINTS1<-HINTS1 %>% 
  mutate (Education = if_else(Educ %in% c("Less than High school", "High school Graduate"),0,
                              if_else (Educ=="Some College",1,
                                       if_else(Educ=="College Graduate or more",2, NULL))))
HINTS1$Education<-factor(HINTS1$Education, levels = c(0:2), labels = c("Less than College","Some College", "College Graduate or more"))

# check
table(HINTS1$Education, HINTS1$Educ,useNA="always")

# race/ethnicity
HINTS1<-HINTS1 %>% 
  mutate (EthRace = if_else(RaceEthnicity == "Non-Hispanic White",0,
                            if_else(RaceEthnicity %in% c("Non-Hispanic Black or African American", "Non-Hispanic Asian","Non-Hispanic Other"),1,
                                                         if_else(RaceEthnicity == "Hispanic",2, NULL))))
HINTS1$EthRace<-factor(HINTS1$EthRace,levels = c(0:2), labels = c("Non-Hispanic White","Non-Hispanic African American/Asian/Other", "Hispanic"))

# check
table(HINTS1$RaceEthnicity,HINTS1$EthRace,useNA="always")
```

# Table 1 after regroup
```{r}
label(HINTS1$Education)<- "Educational level"
label(HINTS1$EthRace)<-"Race & Ethnicity"
table1(~Age+Sex+Education+EthRace+Employment|BreastCan, overall="Total", rowlabelhead="Variable", footnote='SD = standard deviation', HINTS1)
# Those who were diagnosed breast cancer were older than those who weren't diagnosed breast cancer with a mean age of 57.2 (SD = 6.4) and a median age of 59.
```

# univariate logistic regression of association between breastcancer diagnosis and employment
```{r}
model1<-glm(Employment ~ BreastCan, data=HINTS1, family= "binomial")
summary(model1)
odds.n.ends(model1)
```

## Test the linearity assumption for age using the Box Tidwell test.
```{r}
# create term to test linearity
HINTS1<-HINTS1 %>% 
  mutate(age.times.logage = Age*log(Age))

modelagelog<-glm(Employment ~ Age+age.times.logage, data=HINTS1, family= "binomial")
summary(modelagelog)
# p < .05 coefficient means the assumption is violated
```

# Recode age into age group
```{r}
HINTS1<-HINTS1 %>% 
  mutate(agegroup = case_when(Age>=40&Age<45~0,
                              Age>=45&Age<50~1,
                              Age>=50&Age<55~2,
                              Age>=55&Age<60~3,
                              Age>=60&Age<=65~4),
         agegroup = factor(agegroup,0:4, c("Age 40-45","Age 45-50","Age 50-55","Age 55-60","Age 60-65")))

summary(HINTS1$agegroup)
```

# logistic regression with confounders: age, race, education
```{r}
model2<-glm(Employment ~ BreastCan+agegroup+EthRace+Education,data=HINTS1, family= "binomial")
summary(model2)
odds.n.ends(model2)
```

# compare the adjusted model with the original model
```{r}
lrtest(model1, model2)
# significant
```

# Assumptions of multicollinearity using the vif function
```{r}
summ(model2, vifs = TRUE)
# VIF is quite small, the assumptions of multicollinearity met.
```

# look for influential observations using Cook's distance
```{r}
# plot
plot(model2, which = 4, id.n = 3, col="red") 
```

# if exclude influential observations
```{r}
# get Cook's D
model2.data<-augment(model2) %>% 
  mutate(index = 1:n())

# Calculate cutoff and assign to object
cuttoff <- mean(model2.data$.cooksd)*3

# Exclude observations with values greater than the cutoff
HINTS_out<- model2.data %>%
  filter(.cooksd<cuttoff)

# new model without influential observations
model3<-glm(Employment ~ BreastCan+agegroup+EthRace+Education,data=HINTS_out, family= "binomial")
summary(model3)
odds.n.ends(model3)

# compare
compareCoefs(model2, model3)

# sample size is too small, so not exclude influential observations.
```

# evaluate model classification performance 
```{r}
odds.n.ends(model2, predProbPlot=TRUE, rocPlot=TRUE)
```

# effect modification
```{r}
# with interaction term
model4<-glm(Employment ~ BreastCan+agegroup+EthRace+Education+EthRace*BreastCan,data=HINTS1, family= "binomial")
summary(model4)

# compare
lrtest(model2, model4)
# non-significant
```
